---
title: "data_manipulation"
author: "xinyi_lin"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Import FAS csv files

Import litters csv file

```{r}
litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

## Select data

```{r}
select(litters_data, litters_number, everything())   # can reorganize the dataset
```

## Select data practice

In the pups data, select the columns containing litter number, sex, and PD ears.

```{r}
select(pups_data, litter_number, sex, pd_ears)
```

## Filter observation

```{r}
filter(litters_data, gd0_weight < 25)

filter(litters_data, group == "Low8" | group == "Low7")
filter(litters_data, group %in% c("Low7", "Low8"))     #the same


```

Filter to include only pups with sex 1
Filter to include only pups with PD walk less than 11 and sex 2

```{r}
filter(pups_data, sex == 1)
filter(pups_data, pd_walk < 11, sex == 2)
```

## Mutate

Create variables

```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_square = wt_gain^2                   #can use new variables immediately
)
```

## Arrange

Arrange the data...

```{r}
arrange(litters_data, pups_born_alive, gd0_weight)
```

## Pip

Look at interm

```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = mutate(litters_data_selected_cols, 
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
litters_data_with_vars
```

Use nested function call:

```{r}
litters_data_clean = 
  mutate(
    select(
      janitor::clean_names(
        read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
        ), 
    -pups_survive
    ),
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
  )
litters_data_clean
```

Use piping

```{r}
littes_clean_data =                #the name of the finally result
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
  ) %>%
  filter(.data = ., !is.na(gd0_weight)) %>%
  lm(gd18_weight ~ gd0_weight, data = .)     #put the results in the place you want
```

