---
title: "tidy_data"
author: "xinyi_lin"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## gather 

PULSE data

```{r}
pulse_data = haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()
pulse_data
```

This isn't tidy yet....

```{r}
pulse_tidy = pulse_data %>% 
  gather(key = "visit", value = "bdi_score", bdi_score_bl:bdi_score_12m)
```

Illustrate `sepertate`:

```{r}
pulse_tidy %>%
  separate(visit, into = c("bdi_str", "score_str", "visit"), sep = "_") %>%     
  select(-bdi_str, -score_str) %>%
  mutate(visit = replace(visit, visit == "bl", "00m"))
```

## Revisit FAS_litters

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(dose = tolower(dose),
         wt_gain = gd18_weight - gd0_weight) %>%
  arrange(litter_number)
```

## assessment

In the litters data, the variables gd0_weight and gd18_weight give the weight of the mother mouse on gestational days 0 and 18. Write a data cleaning chain that retains only litter_number and these columns; produces new variables gd and weight; and makes gd a numeric variable taking values 0 and 18 (for the last part, you might want to use recode …). Is this version “tidy”?

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>%
  select(litter_number, ends_with("weight")) %>%
  gather(key = "gest_day", value = "weight", gd0_weight:gd18_weight) %>%
  mutate(gest_day, recode(gest_day, "gd0_weight" = 0, "gd18_weight" = 18)) %>%    # replace or add some new col and recode can replace numeric values 
  arrange(litter_number)     # sort tibble
```

## illustrate spread

# create tibble

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)
```

make it readable:

```{r}
analysis_result %>%
  spread(key = time, value = mean) %>%
  knitr::kable()          # make it look likes a table
```

## bind rows

read in LotR data:

```{r}
fellowship_ring = readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring")    # add one col

two_towers = readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
```

create final LotR data:

```{r}
bind_rows(fellowship_ring, two_towers, return_king) %>%
  janitor::clean_names() %>%
  gather(key = sex, value = word, female:male) %>%
  mutate(race = tolower(race))
```

## join data

# preparing data

```{r}
pup_data = read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group))
```

# create joint data

```{r}
FAS_data = left_join(pup_data, litter_data, by = "litter_number")   # left join or right join is not recommended in pipe procedure 
```

